---
title: "Predawn assimilation_plsr"
date: "`r Sys.Date ()`"
output: pdf_document
---

```{r, setup, include=FALSE, cache = F}
knitr::opts_chunk$set(echo = T)
knitr::opts_knit$set(root.dir = '/Users/ting15/Dropbox/Ting_Wang/rice_imaging_ms')
#knitr::opts_knit$set(root.dir = 'C:/Users/rebec/Dropbox/Ting_Wang/rice_imaging_ms')
```

```{r, echo = F, include=FALSE}
rm(list = ls())
## Load package 
list.of.packages <- c("dplyr", "reshape2", "ggplot2", "pls", "scales", "kableExtra", "ggpubr", "gridExtra", "CORElearn", "spectratrait")
invisible(lapply(list.of.packages, library, character.only = TRUE))
```
# Aim  

Predict predawn assimilation (A, $\mu mol m^{-2} s^{-1}$, *n* = 94, N1 = N2 = 47) with hyperspectral data.  

# Methods  

1. Use RReliefFexpRank algorithm to select the top 300 wavelengths that were important for A.   
2. Conduct PCA on the reflectances of the selected wavelengths to concentrate the variance.  
3. Build a PLSR with PCs' scores as input variables.  

```{r, include=FALSE}
## Load datasets  
#HSI: Weekly ID-based data
phyraw <- read.csv('./data/ground_truth/phyraw_forHSI_2022.csv', header = T, sep = ',', dec = '.')
HSI_meta <- read.csv('./data/hyperspectral_imaging/HSI_side_for_biotraits_mod_W6.csv', header = T, sep = ',', dec = '.')
inVar <- "A"
file_path <- "./results/relieF_pca_plsr/A/"

```

```{r, echo = F, fig.align='center', fig.width = 4, fig.height = 4}
trait <- phyraw %>%
  select("ID", "A", "Treatment") %>%
  na.omit() 
#table(trait$Treatment) #N1:47, N2:47
```

## Step 1: Feature selection    

```{r, echo=FALSE, fig.align='center', fig.width = 10, fig.height = 4}
HSI <- HSI_meta[which(HSI_meta$ID %in% trait$ID), -c(746:766)] #drop columns with wv > 2500 and meta data 
wv_trait <- merge(HSI, subset(trait, select = -Treatment), by = "ID")
wv_trait_2 <- wv_trait[ ,-1]

wv_trait_expRank <- attrEval(A ~., wv_trait_2, estimator = "RReliefFexpRank", ReliefIterations = 94*50) 
wv_trait_expRank_long <- data.frame(value = unname(wv_trait_expRank), 
                                       wv = as.numeric(gsub("_", ".", sub("...", "",  names(wv_trait_expRank)))))
wv_trait_expRank_order <- wv_trait_expRank_long[order(wv_trait_expRank_long$value, decreasing = T)[1:300], ] 
wv_trait_expRank_export <- wv_trait_expRank_long[order(wv_trait_expRank_long$value, decreasing = T), ] 
#write.csv(file = paste0(file_path, "wv_trait_expRank.csv"), x= wv_trait_expRank_export, row.names = F)
critical_value <- wv_trait_expRank_long[order(wv_trait_expRank_long$value, decreasing = T)[300], 1]
wv_trait_expRank_long$select <- ifelse(wv_trait_expRank_long$value > critical_value, "Y", "N")

col_attr <- c("Y"="red", "N"="darkgray")
wv_trait_expRank_plot <- ggplot() +
  geom_line(aes(x = wv, y = value), data = wv_trait_expRank_long, col = "black", size = 0.5) +
  scale_x_continuous(breaks = seq(400, 2500, by = 400), labels = seq(400, 2500, by = 400)) +
  geom_point(aes(x = wv, y = value, col = select), data = wv_trait_expRank_long, size = .7,
             show.legend = F) +
  scale_color_manual(values=col_attr) +
  labs(x= "\n Wavelength (nm)", y = "Attribute values") +                     
  theme_bw() +
  theme(
    axis.title = element_text(size = 25, face = "bold"),
    axis.text = element_text(size = 25, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
        )
wv_trait_expRank_plot
```
Figure 1: Attribute values of HSI data on Week 6 for A. Wavelengths with the top 300 attribute values are considered as important (red).  

```{r, include = FALSE}
HSI_feature <- HSI[ , which(as.numeric(gsub("_", ".", sub("...", "",  names(HSI)))) %in% wv_trait_expRank_order$wv)]
HSI_feature2 <- data.frame(ID = HSI$ID, HSI_feature)
HSI_feature2_long <- melt(HSI_feature2, id = "ID")
HSI_feature2_long$variable.num <- as.numeric(gsub("_", ".", sub("...", "", HSI_feature2_long$variable)))
```

```{r, echo = F, eval = F, fig.align='center', fig.width = 10, fig.height = 4}
HSI_p <- ggplot(aes(x=variable.num, y = value, col = ID), data = HSI_feature2_long) +  
  geom_point(size = 0.1) + 
  scale_x_continuous(breaks = seq(400, 2500, by = 200), labels = seq(400, 2500, by = 200)) +
  labs(x = "Wavelength (nm)", y = "Reflectance (%)") + 
  theme_bw() +
 theme(panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "none")
HSI_p
#ggsave(filename = "./results/plsr/A/A_sel_300wv.png", plot = HSI_p, device="png", width = 30, height = 12, units = "cm", dpi = 300)

```

## Step 2: PCA  
The hyperspectral data were not scaled.  

```{r, include=FALSE}
HSI_pca <- prcomp(HSI_feature)
HSI_loadings <- HSI_pca$rotation
HSI_scores <- HSI_pca$x
HSI_sd <- HSI_pca$sdev
HSI_var <- (HSI_pca$sdev)^2
HSI_var_percent <- HSI_var/sum(HSI_var)*100
sum(HSI_var_percent[1:2])
```

```{r, echo = F, warnings = F, fig.align='center', fig.width = 4, fig.height = 4}
#png(filename = "./results/plsr/A/HSI_A_var_percent.png", bg = "white", width = 10, height = 10, units = "in", res = 300)
barplot(HSI_var_percent, xlab='PC', ylab='Percent Variance', names.arg=1:length(HSI_var_percent), las=1, ylim=c(0, max(HSI_var_percent)), col='gray')
#dev.off()
```
Figure 2: PCA scree plot. The first two PCs account for 90.0% of the variance. The top ten loadings in PC1 are wavelengths from 711 to 722 nm. The top ten loadings in PC2 are wavelengths from 1196 to 1349 nm with one wavelength at 758 nm.   
```{r, include=FALSE}
HSI_PC1_loadings <- HSI_loadings[order((HSI_loadings[ ,1])^2, decreasing = T), 1]
HSI_PC2_loadings <- HSI_loadings[order((HSI_loadings[ ,2])^2, decreasing = T), 2]
HSI_top_loadings <- data.frame(Trait = "predawn_QY", PC = rep(c("PC1", "PC2"), each = 10), 
                    ID = c(as.numeric(gsub("_", ".", sub("...", "", names(HSI_PC1_loadings[1:10])))), 
                           as.numeric(gsub("_", ".", sub("...", "", names(HSI_PC2_loadings[1:10]))))
                           ),
                    loadings = c(as.numeric(HSI_PC1_loadings[1:10]),                   
                                 as.numeric(HSI_PC2_loadings[1:10]))
                    )
HSI_top_loadings_2 <- HSI_top_loadings[!duplicated(HSI_top_loadings[ ,c('ID')]), ]
#write.csv(file = "./results/plsr/A/HSI_A_top10_loadings.csv", x= HSI_top_loadings_2, row.names = F)
```

## Step 3: plsr  

```{r, include=FALSE}
`%notin%` <- Negate(`%in%`)
pls::pls.options(plsralg = "oscorespls")
pls::pls.options("plsralg")
opar <- par(no.readonly = T)
```
1. Prepare calibration and evaluation datasets  
The data were first grouped by treatments. Then, 80% of the data in each treatment were sampled for the calibration dataset.  
```{r, echo = F, fig.align='center', fig.width = 9, fig.height = 5}
HSI_I <- as.matrix(HSI_scores)
plsr_data <- data.frame(trait, inputs_scores = I(HSI_I))
set.seed(333)
prop <- 0.8

cal.plsr.data <- plsr_data %>% 
        slice(sample(1:n(), prop*n())) %>% 
        data.frame()
val.plsr.data <- plsr_data[!plsr_data$ID %in% cal.plsr.data$ID,]
print(paste("Cal observations: ",dim(cal.plsr.data)[1],sep="")) #75
print(paste("Val observations: ",dim(val.plsr.data)[1],sep="")) #19

cal_hist_plot <- qplot(cal.plsr.data$A, geom="histogram", bins =12,
                       main = "Cal. (n = 75)",
                       xlab = expression("A ("~mu*mol*m^-2*s^-1~")"), ylab = "Count", 
                       fill=I("#0072B2"), col=I("black")) + theme_bw(base_size = 18)

val_hist_plot <- qplot(val.plsr.data$A, geom="histogram", bins =5,
                       main = "Val. (n = 19)",
                       xlab = expression("A ("~mu*mol*m^-2*s^-1~")"), ylab = "Count",
                       fill=I("#0072B2"),col=I("black")) + theme_bw(base_size = 18)
#val_hist_plot
cal_val_histograms <- grid.arrange(cal_hist_plot, val_hist_plot, ncol=2)
#ggsave(filename = "./results/plsr/A/cal_val_A_histograms.png", plot = cal_val_histograms, device="png", width = 30, height = 12, units = "cm", dpi = 300)
#write.csv(cal.plsr.data, file= "./results/plsr/A/A_Cal_PLSR_Dataset.csv", row.names=FALSE)
#write.csv(val.plsr.data, file="./results/plsr/A/A_Val_PLSR_Dataset.csv", row.names=FALSE)

```
Figure 3: Frequency distributions of A ($\mu mol m^{-2} s^{-1}$) in the calibration (left, *n* = 75) and evaluation (right, *n* = 19) datasets. 
\newpage  

2. Fit plsr model    
Model: A = Scores of PCs. The scores of the PCs were not scaled. 
The model was cross-validated with leave-one-out method.  

```{r, echo = F, fig.align='center', fig.width = 7, fig.height = 5}
plsr.ini <- plsr(A~inputs_scores, scale=FALSE,
                 validation="LOO",trace=FALSE,data=cal.plsr.data)
#png(filename = "./results/plsr/A/A_nComps.png", bg = "white", width = 10, height = 10, units = "in", res = 300)
par(mar = c(6, 7, 4, 2), mgp = c(5, 2, 0))
nComp <- selectNcomp(plsr.ini, method = "onesigma", plot = T,  
                     cex.axis = 2, cex.lab = 2, xlim = c(0, 15), ylim = c(0.4, 0.9),
                     font.lab = 2, font = 2)
#dev.off()
```
Figure 4: Root mean square error of prediction (RMSEP) as a function of model component number in PLSR models. The dotted black vertical line is the number of components that has lowest RMSEP. The vertical long-dashed blue line indicates the number of components suggested. The points and error bars are the means and standard errors of RMSEP in the cross validation, respectively.  