---
title: "E_plsr"
date: "`r Sys.Date ()`"
output: pdf_document
---

```{r, setup, include=FALSE, cache = F}
knitr::opts_chunk$set(echo = T)
knitr::opts_knit$set(root.dir = '/Users/ting15/Dropbox/Ting_Wang/rice_imaging_ms')
#knitr::opts_knit$set(root.dir = 'C:/Users/rebec/Dropbox/Ting_Wang/rice_imaging_ms')
```

```{r, echo = F, include=FALSE}
rm(list = ls())
## Load package 
list.of.packages <- c("dplyr", "reshape2", "ggplot2", "pls", "scales", "kableExtra", "ggpubr", "gridExtra", "CORElearn", "spectratrait")
invisible(lapply(list.of.packages, library, character.only = TRUE))
```
# Aim  

Predict E ($molm^{-2}s^{-1}$, *n* = 94, N1 = 47, N2 = 47) with hyperspectral data.  

# Methods  

1. Use RReliefFexpRank algorithm to select the top 300 wavelengths that were important for *E* ($molm^{-2}s^{-1}$).    
2. Conduct PCA on the selected wavelengths to concentrate the variance.  
3. Build a PLSR with PCs' scores as input variables.  

```{r, include=FALSE}
## Load datasets  
#HSI: Weekly ID-based data
phyraw <- read.csv('./data/ground_truth/phyraw_forHSI_2022.csv', header = T, sep = ',', dec = '.')
HSI_meta <- read.csv('./data/hyperspectral_imaging/HSI_side_for_biotraits_mod_W6.csv', header = T, sep = ',', dec = '.')
inVar <- "E"
file_path <- "./results/relieF_pca_plsr/E/"
```

```{r, echo = F, fig.align='center', fig.width = 4, fig.height = 4}
trait <- phyraw %>%
  select("ID", "E", "Treatment") %>%
  na.omit() 

#table(trait$Treatment) #N1: 47, N2: 47
```

## Step 1: Feature selection    

```{r, echo=FALSE, fig.align='center', fig.width = 10, fig.height = 4}
HSI <- HSI_meta[which(HSI_meta$ID %in% trait$ID), -c(746:766)] #drop columns with wv > 2500 and meta data 
wv_trait <- merge(HSI, subset(trait, select = -c(Treatment)), by = "ID")
wv_trait_2 <- wv_trait[ ,-1]

wv_trait_expRank <- attrEval(E ~., wv_trait_2, estimator = "RReliefFexpRank", ReliefIterations = nrow(trait)*50) 
wv_trait_expRank_long <- data.frame(value = unname(wv_trait_expRank), 
                                       wv = as.numeric(gsub("_", ".", sub("...", "",  names(wv_trait_expRank)))))
critical_value <- wv_trait_expRank_long[order(wv_trait_expRank_long$value, decreasing = T)[300], 1]
wv_trait_expRank_long$select <- ifelse(wv_trait_expRank_long$value > critical_value, "Y", "N")

wv_trait_expRank_order <- wv_trait_expRank_long[order(wv_trait_expRank_long$value, decreasing = T)[1:300], ] 

wv_trait_expRank_export <- wv_trait_expRank_long[order(wv_trait_expRank_long$value, decreasing = T), ] 
#write.csv(file = paste0(file_path, "wv_trait_expRank.csv"), x= wv_trait_expRank_export, row.names = F)
col_attr <- c("Y"="red", "N"="darkgray")
wv_trait_expRank_plot <- ggplot() +
  geom_line(aes(x = wv, y = value), data = wv_trait_expRank_long, col = "black", size = 0.5) +
  scale_x_continuous(breaks = seq(400, 2500, by = 400), labels = seq(400, 2500, by = 400)) +
  scale_y_continuous(limits = c(-0.07, 0.08)) +
  geom_point(aes(x = wv, y = value, col = select), data = wv_trait_expRank_long, size = .7,
             show.legend = F) +
  scale_color_manual(values=col_attr) +
  labs(x= "\n Wavelength (nm)", y = "Attribute values") +                     
  theme_bw() +
  theme(
    axis.title = element_text(size = 25, face = "bold"),
    axis.text = element_text(size = 25, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
        )
wv_trait_expRank_plot
```
Figure 1: Attribute values of HSI data on Week 6 for *E*. Wavelengths with the top 300 attribute values are considered as important (red). 
```{r, include = FALSE}
HSI_feature <- HSI[ , which(as.numeric(gsub("_", ".", sub("...", "",  names(HSI)))) %in% wv_trait_expRank_order$wv)]
HSI_feature2 <- data.frame(ID = HSI$ID, HSI_feature)
HSI_feature2_long <- melt(HSI_feature2, id = "ID")
HSI_feature2_long$variable.num <- as.numeric(gsub("_", ".", sub("...", "", HSI_feature2_long$variable)))
```

```{r, echo = F, eval = F, fig.align='center', fig.width = 10, fig.height = 4}
HSI_p <- ggplot(aes(x=variable.num, y = value, col = ID), data = HSI_feature2_long) +  
  geom_point(size = 0.1) + 
  scale_x_continuous(breaks = seq(400, 2500, by = 200), labels = seq(400, 2500, by = 200)) +
  labs(x = "Wavelength (nm)", y = "Reflectance (%)") + 
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position = "none")
HSI_p
#ggsave(filename = paste0(file_path, "E_sel_300wv.png"), plot = HSI_p, device="png", width = 30, height = 12, units = "cm", dpi = 300)

```

## Step 2: PCA  
The hyperspectral data were not scaled.  

```{r, include=FALSE}
HSI_pca <- prcomp(HSI_feature)
HSI_loadings <- HSI_pca$rotation
HSI_scores <- HSI_pca$x
HSI_sd <- HSI_pca$sdev
HSI_var <- (HSI_pca$sdev)^2
HSI_var_percent <- HSI_var/sum(HSI_var)*100
sum(HSI_var_percent[1:3]) #94.8
```

```{r, echo = F, warnings = F, fig.align='center', fig.width = 4, fig.height = 4}
#png(filename = paste0(file_path, "HSI_E_var_percent.png"), bg = "white", width = 10, height = 10, units = "in", res = 300)
barplot(HSI_var_percent, xlab='PC', ylab='Percent Variance', names.arg=1:length(HSI_var_percent), las=1, ylim=c(0, max(HSI_var_percent)), col='gray')
#dev.off()
```
Figure 2: PCA scree plot. The first three PCs account for 94.8% of the variance. The top ten loadings in PC1, PC2 and PC3 are in the regions of 711 to 721 nm, 1140 to 1292 nm and 2284 to 2497 nm, respectively.
```{r, include=FALSE}
HSI_PC1_loadings <- HSI_loadings[order((HSI_loadings[ ,1])^2, decreasing = T), 1]
HSI_PC2_loadings <- HSI_loadings[order((HSI_loadings[ ,2])^2, decreasing = T), 2]
HSI_PC3_loadings <- HSI_loadings[order((HSI_loadings[ ,3])^2, decreasing = T), 3]

HSI_top10_loadings <- data.frame(Trait = "E", PC = rep(c("PC1", "PC2", "PC3"), each = 10), 
                    ID = c(as.numeric(gsub("_", ".", sub("...", "", names(HSI_PC1_loadings[1:10])))), 
                           as.numeric(gsub("_", ".", sub("...", "", names(HSI_PC2_loadings[1:10])))),
                           as.numeric(gsub("_", ".", sub("...", "", names(HSI_PC3_loadings[1:10]))))
                           ),
                    loadings = c(as.numeric(HSI_PC1_loadings[1:10]),                   
                                 as.numeric(HSI_PC2_loadings[1:10]), 
                                 as.numeric(HSI_PC3_loadings[1:10])))
HSI_top10_loadings_2 <- HSI_top10_loadings[!duplicated(HSI_top10_loadings[ ,c('ID')]), ]
#write.csv(file = paste0(file_path, "HSI_E_top_loadings.csv"), x= HSI_top10_loadings_2, row.names = F)
```

## Step 3: plsr  
 
```{r, include=FALSE}
`%notin%` <- Negate(`%in%`)
pls::pls.options(plsralg = "oscorespls")
pls::pls.options("plsralg")
opar <- par(no.readonly = T)
```
1. Prepare calibration and evaluation datasets  
The data were first grouped by treatment. Then, 80% of the data in each treatment level were sampled for the calibration dataset.
```{r, echo = F, fig.align='center', fig.width = 9, fig.height = 5}
HSI_I <- as.matrix(HSI_scores)
plsr_data <- data.frame(trait, inputs_scores = I(HSI_I))
set.seed(5678)
prop <- 0.8
group_variables <- c("Treatment")
cal.plsr.data <- plsr_data %>% 
        group_by_at(vars(all_of(group_variables))) %>%
        slice(sample(1:n(), prop*n())) %>% 
        data.frame()
val.plsr.data <- plsr_data[!plsr_data$ID %in% cal.plsr.data$ID,]
print(paste("Cal observations: ",dim(cal.plsr.data)[1],sep="")) #74
print(paste("Val observations: ",dim(val.plsr.data)[1],sep="")) #20

cal_hist_plot <- qplot(cal.plsr.data$E, geom="histogram", bins =10,
                       main = "Cal. (n = 74)",
                       xlab = expression("E ("*mol*m^-2*s^-1*")"), ylab = "Count", 
                       fill=I("#0072B2"), col=I("black")) + theme_bw(base_size = 18) +
  theme(axis.text.x = element_text(hjust = 1))

val_hist_plot <- qplot(val.plsr.data$E, geom="histogram", bins =5,
                       main = "Val. (n = 20)",
                       xlab = expression("E ("*mol*m^-2*s^-1*")"), ylab = "Count", 
                       fill=I("#0072B2"), col=I("black")) + theme_bw(base_size = 18)
cal_val_histograms <- grid.arrange(cal_hist_plot, val_hist_plot, ncol=2)
#ggsave(filename = paste0(file_path, "cal_val_E_histograms.png"), plot = cal_val_histograms, device="png", width = 30, height = 12, units = "cm", dpi = 300)
#write.csv(cal.plsr.data, file = paste0(file_path, "E_Cal_PLSR_Dataset.csv"), row.names=FALSE)
#write.csv(val.plsr.data, file = paste0(file_path, "E_Val_PLSR_Dataset.csv"), row.names=FALSE)

```
Figure 3: Frequency distributions of E ($molm^{-2}s^{-1}$) in the calibration (*n* = 74) and evaluation (*n* = 20) datasets. 
\newpage  

2. Fit a plsr model    
Model: E = Scores of PCs. The scores of the PCs were not scaled. 
The model was cross-validated with leave-one-out method.  
```{r, echo = F, fig.align='center', fig.width = 9, fig.height = 5}
plsr.ini <- plsr(E~inputs_scores, scale=FALSE,
                 validation="LOO",trace=FALSE,data=cal.plsr.data)
#png(filename = paste0(file_path, "E_nComps.png"), bg = "white", width = 10, height = 10, units = "in", res = 300)
par(mar = c(6, 7, 4, 2), mgp = c(5, 2, 0))
nComp <- selectNcomp(plsr.ini, method = "onesigma", plot = T, 
                     cex.axis = 2, cex.lab = 2, xlim = c(0, 15), ylim = c(0.001, 0.002), font.lab = 2, font = 2)
#dev.off()
```
Figure 4: Root mean square error of prediction (RMSEP) as a function of model component number in PLSR models. The dotted black vertical line is the number of components that has the lowest RMSEP. The vertical long-dashed blue line is the suggested number of components. The points and error bars are the means and standard errors of RMSEP in the cross validation, respectively.  