---
title: "final biomass_whole_spectrum"
date: "`r Sys.Date ()`"
output:
  pdf_document: default
---

```{r, setup, include=FALSE, cache = F}
knitr::opts_chunk$set(echo = T)
knitr::opts_knit$set(root.dir = '/Users/ting15/Dropbox/Ting_Wang/rice_imaging_ms')
#knitr::opts_knit$set(root.dir = 'C:/Users/rebec/Dropbox/Ting_Wang/rice_imaging_ms')
```

```{r, echo = F, include=FALSE}
rm(list = ls())
## Load package 
list.of.packages <- c("dplyr", "reshape2", "ggplot2", "pls", "scales", "kableExtra", "ggpubr", "gridExtra", "CORElearn", "spectratrait")
invisible(lapply(list.of.packages, library, character.only = TRUE))
```
# Aim  

Predict final biomass (g) (*n* = 94, N1:47, N2:47) with hyperspectral data.  
   
# Methods  

1. Use RReliefFexpRank algorithm to select the top 300 wavelengths that were important for final biomass.  
2. Conduct PCA on the reflectances of the selected wavelengths to concentrate the variance.  
3. Build a PLSR with PCs' scores as input variables.    

```{r, include=FALSE}
## Load datasets  
#HSI: Weekly ID-based data
phyraw <- read.csv('./data/ground_truth/phyraw_forHSI_2022.csv', header = T, sep = ',', dec = '.')
HSI_meta <- read.csv('./data/hyperspectral_imaging/HSI_side_for_biotraits_mod_W13.csv', header = T, sep = ',', dec = '.')
inVar <- "FB"
file_path <- "./results/relieF_pca_plsr/final_biomass/"
```

```{r, echo=FALSE, fig.align='center', fig.width = 4, fig.height = 4}
trait <- phyraw %>%
  select("ID", all_of(inVar), "Treatment") %>%
  na.omit() 
#table(trait$Treatment)
```

## Step 1: Feature selection  

```{r, echo=FALSE, fig.align='center', fig.width = 10, fig.height = 4}
HSI <- HSI_meta[which(HSI_meta$ID %in% trait$ID), -c(746:766)] #drop columns with wv > 2500 and meta data 

wv_trait <- merge(HSI, subset(trait, select = -Treatment), by = "ID")
wv_trait_2 <- wv_trait[ ,-1]

wv_trait_expRank <- attrEval(FB ~., wv_trait_2, estimator = "RReliefFexpRank", ReliefIterations = 94*50) 
wv_trait_expRank_long <- data.frame(value = unname(wv_trait_expRank), 
                                       wv = as.numeric(gsub("_", ".", sub("...", "",  names(wv_trait_expRank)))))
critical_value <- wv_trait_expRank_long[order(wv_trait_expRank_long$value, decreasing = T)[300], 1]
wv_trait_expRank_long$select <- ifelse(wv_trait_expRank_long$value > critical_value, "Y", "N")
wv_trait_expRank_order <- wv_trait_expRank_long[order(wv_trait_expRank_long$value, decreasing = T)[1:300], ]

wv_trait_expRank_export <- wv_trait_expRank_long[order(wv_trait_expRank_long$value, decreasing = T), ] 
#write.csv(file = paste0(file_path, "wv_trait_expRank.csv"), x= wv_trait_expRank_export, row.names = F)

col_attr <- c("Y"="red", "N"="darkgray")
wv_trait_expRank_plot <- ggplot() +
  geom_line(aes(x = wv, y = value), data = wv_trait_expRank_long, col = "black", size = 0.5) +
  scale_x_continuous(breaks = seq(400, 2500, by = 400), labels = seq(400, 2500, by = 400)) +
  scale_y_continuous(breaks = seq(-0.1, 0.2, by = 0.1), labels = seq(-0.1, 0.2, by = 0.1),
                     limits = c(-0.13, 0.23)) +
  geom_point(aes(x = wv, y = value, col = select), data = wv_trait_expRank_long, size = .7,
             show.legend = F) +
  scale_color_manual(values=col_attr) +
  labs(x= "\n Wavelength (nm)", y = "Attribute values") +                     
  theme_bw() +
  theme(
    axis.title = element_text(size = 25, face = "bold"),
    axis.text = element_text(size = 25, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
        )
wv_trait_expRank_plot
```

```{r, echo=FALSE, eval=FALSE}
ggsave(filename = paste0(file_path, "final_biomass_expRank_no_h_line.png"), plot = wv_trait_expRank_plot, 
device="png", width = 30, height = 12, units = "cm", dpi = 400)
```
Figure 1: Attribute values of HSI data on Week 13 for final biomass. Wavelengths with the top 300 attribute values are considered as important (red).  
```{r, include = FALSE}
HSI_feature <- HSI[ , which(as.numeric(gsub("_", ".", sub("...", "",  names(HSI)))) %in% wv_trait_expRank_order$wv)]
head(HSI_feature)
HSI_feature2 <- data.frame(ID = HSI$ID, HSI_feature)
HSI_feature2_long <- melt(HSI_feature2, id = "ID")
HSI_feature2_long$variable.num <- as.numeric(gsub("_", ".", sub("...", "", HSI_feature2_long$variable)))
```

```{r, echo = F, eval = F, fig.align='center', fig.width = 10, fig.height = 4}
HSI_p <- ggplot(aes(x=variable.num, y = value, col = ID), data = HSI_feature2_long) +  
  geom_point(size = 0.1) + 
  scale_x_continuous(breaks = seq(400, 2500, by = 200), labels = seq(400, 2500, by = 200)) +
  labs(x = "Wavelength (nm)", y = "Reflectance (%)") + 
  theme_bw() +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "none")
HSI_p
#ggsave(filename = paste0(file_path, "final_biomass_sel_300wv.png"), plot = HSI_p, device="png", width = 30, height = 12, units = "cm", dpi = 300)

``` 

## Step 2: PCA  
Note that the HSI data were not scaled.  

```{r, include=FALSE}
HSI_pca <- prcomp(HSI_feature)
HSI_loadings <- HSI_pca$rotation
HSI_scores <- HSI_pca$x
HSI_sd <- HSI_pca$sdev
HSI_var <- (HSI_pca$sdev)^2
HSI_var_percent <- HSI_var/sum(HSI_var)*100
sum(HSI_var_percent[1:3])
```

```{r, echo = F, warnings = F, fig.align='center', fig.width = 4, fig.height = 4}
#png(filename = "./results/plsr/final_biomass/HSI_final_biomass_var_percent.png", bg = "white", width = 10, height = 10, units = "in", res = 300)
barplot(HSI_var_percent, xlab='PC', ylab='Percent Variance', names.arg=1:length(HSI_var_percent), las=1, ylim=c(0, max(HSI_var_percent)), col='gray')
#dev.off()
```
Figure 2: PCA scree plot. PCs 1-3 account for 94.5% of the variance. The top ten loadings in PC1 are wavelengths located in regions of 1405 to 1422 nm and 1860 to 1899 nm. In PC2, the top ten loadings are wavelengths from 715 to 726 nm. In PC3, the top ten loadings are wavelengths mainly from 742 to 752 nm, with one wavelength at 1708 nm.    

```{r, include=FALSE}
HSI_PC1_loadings <- HSI_loadings[order((HSI_loadings[ ,1])^2, decreasing = T), 1]
HSI_PC2_loadings <- HSI_loadings[order((HSI_loadings[ ,2])^2, decreasing = T), 2]
HSI_PC3_loadings <- HSI_loadings[order((HSI_loadings[ ,3])^2, decreasing = T), 3]
HSI_top10_loadings <- data.frame(Trait = "final_biomass", PC = rep(c("PC1", "PC2", "PC3"), each = 10), 
                    ID = c(as.numeric(gsub("_", ".", sub("...", "", names(HSI_PC1_loadings[1:10])))), 
                           as.numeric(gsub("_", ".", sub("...", "", names(HSI_PC2_loadings[1:10])))),
                           as.numeric(gsub("_", ".", sub("...", "", names(HSI_PC3_loadings[1:10]))))
                           ),
                    loadings = c(as.numeric(HSI_PC1_loadings[1:10]),                   
                                 as.numeric(HSI_PC2_loadings[1:10]), 
                                 as.numeric(HSI_PC3_loadings[1:10])))
HSI_top10_loadings_2 <- HSI_top10_loadings[!duplicated(HSI_top10_loadings[ ,c('ID')]), ]
#write.csv(file = paste0(file_path, "HSI_final_biomass_top10_loadings_2.csv"), x= HSI_top10_loadings_2, row.names = F)
```

## Step 3: plsr  

```{r, include=FALSE}
`%notin%` <- Negate(`%in%`)
pls::pls.options(plsralg = "oscorespls")
pls::pls.options("plsralg")
opar <- par(no.readonly = T)
```
1. Prepare calibration and evaluation datasets  
The data were first grouped by treatment. Then, 80% of the data in each treatment level were sampled for the calibration dataset.
```{r, echo = F, fig.align='center', fig.width = 9, fig.height = 5}
HSI_I <- as.matrix(HSI_scores)
plsr_data <- data.frame(trait, inputs_scores = I(HSI_I))
set.seed(14)
prop <- 0.8
group_variables <- c("Treatment")
cal.plsr.data <- plsr_data %>% 
        group_by_at(vars(all_of(group_variables))) %>%
        slice(sample(1:n(), prop*n())) %>% 
        data.frame()
val.plsr.data <- plsr_data[!plsr_data$ID %in% cal.plsr.data$ID,]
print(paste("Cal observations: ",dim(cal.plsr.data)[1],sep="")) #74
print(paste("Val observations: ",dim(val.plsr.data)[1],sep="")) #20

cal_hist_plot <- qplot(cal.plsr.data$FB, geom="histogram", binwidth =30,
                       main = "Cal. (n = 74)",
                       xlab = "Final biomass (g)", ylab = "Count", 
                       fill=I("#0072B2"), col=I("black")) + theme_bw(base_size = 18)
val_hist_plot <- qplot(val.plsr.data$FB, geom="histogram", binwidth =30,
                       main = "Val. (n = 20)",
                       xlab = "Final biomass (g)", ylab = "Count", 
                       fill=I("#0072B2"),col=I("black")) + theme_bw(base_size = 18)
cal_val_histograms <- grid.arrange(cal_hist_plot, val_hist_plot, ncol=2)
```
Figure 3: Frequency distributions of final biomass (g) in the calibration (left, *n* = 74) and evaluation (right, *n* = 20) datasets.  
```{r, echo=FALSE, eval=FALSE, fig.align='center', fig.width = 5, fig.height = 5 }
cal_hist_plot
ggsave(filename = paste0(file_path, "cal_fb_histogram.png"), plot = cal_hist_plot, device="png", width = 12, height = 12, units = "cm", dpi = 300)
val_hist_plot
ggsave(filename = paste0(file_path, "val_fb_histogram.png"), plot = val_hist_plot, device="png", width = 12, height = 12, units = "cm", dpi = 300)
#ggsave(filename = paste0(file_path, "cal_val_final_biomass_histograms.png"), plot = cal_val_histograms, device="png", width = 30, height = 12, units = "cm", dpi = 300)
#write.csv(x = cal.plsr.data, file= paste0(file_path, "final_biomass_Cal_PLSR_Dataset.csv"), row.names=FALSE)
#write.csv(x = val.plsr.data, file= paste0(file_path, "final_biomass_Val_PLSR_Dataset.csv"), row.names=FALSE)

```
 
2. Fit a plsr model  
Model: final biomass = Scores of PCs. The scores of the PCs were not scaled.  
The model was cross-validated with leave-one-out method.   

```{r, echo = F, fig.align='center', fig.width = 7, fig.height = 5}
plsr.ini <- plsr(FB~inputs_scores, scale=FALSE,
                 validation="LOO",trace=FALSE,data=cal.plsr.data)
#png(filename = "./results/plsr/final_biomass/final_biomass_nComps.png", bg = "white", width = 10, height = 10, units = "in", res = 300)
par(mar = c(6, 7, 4, 2), mgp = c(5, 2, 0))
nComp <- selectNcomp(plsr.ini, method = "onesigma", plot = T, 
                     cex.axis = 2, cex.lab = 2, xlim = c(0, 15), ylim = c(20, 34),
                     font.lab = 2, font = 2)
#dev.off()
```
Figure 4: Root mean square error of prediction (RMSEP) as a function of model component number in PLSR models. The dotted black vertical line is the number of components that has the lowest RMSEP. The vertical long-dashed blue line indicates the suggested number of components. The points and error bars are the means and standard errors of RMSEP in the cross validation, respectively.