---
title: "midday QY_plsr"
date: "`r Sys.Date ()`"
output: pdf_document
---

```{r, setup, include=FALSE, cache = F}
knitr::opts_chunk$set(echo = T)
knitr::opts_knit$set(root.dir = '/Users/ting15/Dropbox/Ting_Wang/rice_imaging_ms')
#knitr::opts_knit$set(root.dir = 'C:/Users/rebec/Dropbox/Ting_Wang/rice_imaging_ms')
```

```{r, echo = F, include=FALSE}
rm(list = ls())
## Load package 
list.of.packages <- c("dplyr", "reshape2", "ggplot2", "pls", "scales", "kableExtra", "ggpubr", "gridExtra", "CORElearn", "spectratrait")
invisible(lapply(list.of.packages, library, character.only = TRUE))
```
# Aim  

Predict midday QY (mQY, *n* = 94, N1=N2=47) with hyperspectral data.  

# Methods  

1. Use RReliefFexpRank algorithm to select the top 300 wavelengths that were important for mQY.  
2. Conduct PCA on the reflectances of the selected wavelengths to concentrate the variance.  
3. Build a PLSR with PCs' scores as input variables.   

```{r, include=FALSE}
## Load datasets  
#HSI: Weekly ID-based data
phyraw <- read.csv('./data/ground_truth/phyraw_forHSI_2022.csv', header = T, sep = ',', dec = '.')
HSI_meta <- read.csv('./data/hyperspectral_imaging/HSI_side_for_biotraits_mod_W10.csv', header = T, sep = ',', dec = '.')
inVar <- "mQY"
file_path <- "./results/relieF_pca_plsr/midday_QY/"
```

```{r, echo = F, fig.align='center', fig.width = 4, fig.height = 4}
trait <- phyraw %>%
  select("ID", all_of(inVar), "Treatment") %>%
  na.omit() 
#table(trait$Treatment) #N1:47, N2:47
```

## Step 1: Feature selection    

```{r, echo=FALSE, fig.align='center', fig.width = 10, fig.height = 4}
HSI <- HSI_meta[which(HSI_meta$ID %in% trait$ID), -c(746:766)] #drop columns with wv > 2500 and meta data 
wv_trait <- merge(HSI, subset(trait, select = -c(Treatment)), by = "ID")
wv_trait_2 <- wv_trait[ ,-1]

wv_trait_expRank <- attrEval(mQY ~., wv_trait_2, estimator = "RReliefFexpRank", ReliefIterations = nrow(trait)*50) 
wv_trait_expRank_long <- data.frame(value = unname(wv_trait_expRank), 
                                       wv = as.numeric(gsub("_", ".", sub("...", "",  names(wv_trait_expRank)))))
critical_value <- wv_trait_expRank_long[order(wv_trait_expRank_long$value, decreasing = T)[300], 1]
wv_trait_expRank_long$select <- ifelse(wv_trait_expRank_long$value > critical_value, "Y", "N")
wv_trait_expRank_order <- wv_trait_expRank_long[order(wv_trait_expRank_long$value, decreasing = T)[1:300], ] 
wv_trait_expRank_export <- wv_trait_expRank_long[order(wv_trait_expRank_long$value, decreasing = T), ] 
#write.csv(file = paste0(file_path, "wv_trait_expRank.csv"), x= wv_trait_expRank_export, row.names = F)
col_attr <- c("Y"="red", "N"="darkgray")
wv_trait_expRank_plot <- ggplot() +
  geom_line(aes(x = wv, y = value), data = wv_trait_expRank_long, col = "black", size = 0.5) +
  scale_x_continuous(breaks = seq(400, 2500, by = 400), labels = seq(400, 2500, by = 400)) +
  geom_point(aes(x = wv, y = value, col = select), data = wv_trait_expRank_long, size = .7,
             show.legend = F) +
  scale_color_manual(values=col_attr) +
  labs(x= "\n Wavelength (nm)", y = "Attribute values") +                     
  theme_bw() +
  theme(
    axis.title = element_text(size = 25, face = "bold"),
    axis.text = element_text(size = 25, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
        )
wv_trait_expRank_plot
```
Figure 1: Attribute values of HSI data on Week 10 for mQY. Wavelengths with the top 300 attribute values are considered as important (red).      
```{r, include = FALSE}
HSI_feature <- HSI[ , which(as.numeric(gsub("_", ".", sub("...", "",  names(HSI)))) %in% wv_trait_expRank_order$wv)]
HSI_feature2 <- data.frame(ID = HSI$ID, HSI_feature)
HSI_feature2_long <- melt(HSI_feature2, id = "ID")
HSI_feature2_long$variable.num <- as.numeric(gsub("_", ".", sub("...", "", HSI_feature2_long$variable)))
```

```{r, echo = F, eval = F, fig.align='center', fig.width = 7, fig.height = 4}
HSI_p <- ggplot(aes(x=variable.num, y = value, col = ID), data = HSI_feature2_long) +  
  geom_point(size = 0.1) + 
  scale_x_continuous(breaks = seq(400, 2500, by = 200), labels = seq(400, 2500, by = 200)) +
  labs(x = "Wavelength (nm)", y = "Reflectance (%)") + 
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.text.x = element_text(size = 10, hjust = 1),
        legend.position = "none")
HSI_p
#ggsave(filename = paste0(file_path, "/midday_QY_sel_300wv.png"), plot = HSI_p, device="png", width = 30, height = 12, units = "cm", dpi = 300)

```

## Step 2: PCA  
The hyperspectral data were not scaled.  

```{r, include=FALSE}
HSI_pca <- prcomp(HSI_feature)
HSI_loadings <- HSI_pca$rotation
HSI_scores <- HSI_pca$x
HSI_sd <- HSI_pca$sdev
HSI_var <- (HSI_pca$sdev)^2
HSI_var_percent <- HSI_var/sum(HSI_var)*100
sum(HSI_var_percent[1:3]) #90.5
```

```{r, echo = F, warnings = F, fig.align='center', fig.width = 4, fig.height = 4}
#png(filename = paste0(file_path, "/HSI_midday_QY_var_percent.png"), bg = "white", width = 10, height = 10, units = "in", res = 300)
barplot(HSI_var_percent, xlab='PC', ylab='Percent Variance', names.arg=1:length(HSI_var_percent), las=1, ylim=c(0, max(HSI_var_percent)), col='gray')
#dev.off()
```
Figure 2: PCA scree plot. The first three PCs account for 90.5% of the variance. The top ten loadings in PC1, PC2 and PC3 are wavelengths from 712 to 724 nm, from 981 to 1066 nm and from 2408 to 2497 nm, respectively.  

```{r, include=FALSE}
HSI_PC1_loadings <- HSI_loadings[order((HSI_loadings[ ,1])^2, decreasing = T), 1]
HSI_PC2_loadings <- HSI_loadings[order((HSI_loadings[ ,2])^2, decreasing = T), 2]
HSI_PC3_loadings <- HSI_loadings[order((HSI_loadings[ ,3])^2, decreasing = T), 3]

HSI_top_loadings <- data.frame(Trait = "midday_QY", PC = rep(c("PC1", "PC2", "PC3"), each = 10), 
                    ID = c(as.numeric(gsub("_", ".", sub("...", "", names(HSI_PC1_loadings[1:10])))), 
                           as.numeric(gsub("_", ".", sub("...", "", names(HSI_PC2_loadings[1:10])))),
                           as.numeric(gsub("_", ".", sub("...", "", names(HSI_PC3_loadings[1:10]))))
                           ),
                    loadings = c(as.numeric(HSI_PC1_loadings[1:10]),                   
                                 as.numeric(HSI_PC2_loadings[1:10]),
                                 as.numeric(HSI_PC3_loadings[1:10]))
                    )
HSI_top_loadings_2 <- HSI_top_loadings[!duplicated(HSI_top_loadings[ ,c('ID')]), ]
#write.csv(file = paste0(file_path, "/HSI_midday_QY_top_loadings.csv"), x= HSI_top_loadings_2, row.names = F)
```

## Step 3: plsr  

```{r, include=FALSE}
`%notin%` <- Negate(`%in%`)
pls::pls.options(plsralg = "oscorespls")
pls::pls.options("plsralg")
opar <- par(no.readonly = T)
```
1. Prepare calibration and evaluation datasets  
The data were first grouped by treatment. Then, 80% of the data were sampled from each treatment level for the calibration dataset.
```{r, echo = F, fig.align='center', fig.width = 9, fig.height = 5}
HSI_I <- as.matrix(HSI_scores)
plsr_data <- data.frame(trait, inputs_scores = I(HSI_I))
set.seed(1212)
prop <- 0.8
group_variables <- c("Treatment")
cal.plsr.data <- plsr_data %>% 
        group_by_at(vars(all_of(group_variables))) %>%
        slice(sample(1:n(), prop*n())) %>% 
        data.frame()
val.plsr.data <- plsr_data[!plsr_data$ID %in% cal.plsr.data$ID,]
print(paste("Cal observations: ",dim(cal.plsr.data)[1],sep="")) #74
print(paste("Val observations: ",dim(val.plsr.data)[1],sep="")) #20

cal_hist_plot <- qplot(cal.plsr.data$mQY, geom="histogram", bins =7,
                       main = "Cal. (n = 74)",
                       xlab = "midday QY", ylab = "Count", 
                       fill=I("#0072B2"), col=I("black")) + theme_bw(base_size = 18)

val_hist_plot <- qplot(val.plsr.data$mQY, geom="histogram", bins =4,
                       main = "Val. (n = 20)",
                       xlab = "midday QY", ylab = "Count", 
                       fill=I("#0072B2"),col=I("black")) + theme_bw(base_size = 18)
#val_hist_plot
cal_val_histograms <- grid.arrange(cal_hist_plot, val_hist_plot, ncol=2)
#ggsave(filename = paste0(file_path, "cal_val_midday_QY_histograms.png"), plot = cal_val_histograms, device="png", width = 30, height = 12, units = "cm", dpi = 300)
#write.csv(cal.plsr.data, file = paste0(file_path, "midday_QY_Cal_PLSR_Dataset.csv"), row.names=FALSE)
#write.csv(val.plsr.data, file = paste0(file_path, "midday_QY_Val_PLSR_Dataset.csv"), row.names=FALSE)

```
Figure 3: Frequency distributions of mQY in the calibration (left, *n* = 74) and evaluation (right, *n* = 20) datasets. 

2. Fit a plsr model    
Model: mQY = Scores of PCs. The scores of the PCs were not scaled. 
The model was cross-validated with leave-one-out method.  
```{r, echo = F, fig.align='center', fig.width = 9, fig.height = 7}
plsr.ini <- plsr(mQY~inputs_scores, scale=FALSE,
                 validation="LOO",trace=FALSE,data=cal.plsr.data)
#png(filename = paste0(file_path, "midday_QY_nComps.png"), bg = "white", width = 10, height = 10, units = "in", res = 300)
par(mar = c(6, 7, 4, 2), mgp = c(5, 2, 0))
nComp <- selectNcomp(plsr.ini, method = "onesigma", plot = T, 
                     cex.axis = 2, cex.lab = 2, xlim = c(0, 15), ylim = c(0.02, 0.06),
                     font.lab = 2, font = 2)
#dev.off()
```
Figure 4: Root mean square error of prediction (RMSEP) as a function of model component number in PLSR models. The dotted black vertical line is the number of components that has the lowest RMSEP. The vertical long-dashed blue line indicates the suggested number of components. The points and error bars are the mean and standard errors of RMSEP in the cross validation, respectively.  