---
title: "N_reliefF_PCA_plsr"
date: "`r Sys.Date ()`"
output:
  pdf_document: default

---
  

```{r, setup, include=FALSE, cache = F}
# Set directory
knitr::opts_chunk$set(echo = T)
knitr::opts_knit$set(root.dir = '/Users/ting15/Dropbox/Ting_Wang/rice_imaging_ms')
# Readers please set their own directory.
```

```{r, echo = F, include=FALSE}
rm(list = ls())
# Load packages
list.of.packages <- c("dplyr", "reshape2", "ggplot2", "pls", "scales", "kableExtra", "ggpubr", "gridExtra", "CORElearn", "spectratrait")
invisible(lapply(list.of.packages, library, character.only = TRUE))
```
# Aim  

Quantify nitrogen (%, N) with hyperspectral imaging data (HSI data).  

# Methods  

1. Use RReliefFexpRank algorithm to select the top 300 wavelengths that were important for N (%).  
2. Conduct PCA on the reflectances of the selected wavelengths to concentrate the variance.  
3. Build a PLSR model with PCs' scores as input variables.  

```{r, echo = F, include=FALSE}
phyraw <- read.csv('./data/ground_truth/phyraw_forHSI_2022.csv', header = T, sep = ',', dec = '.')
ref_meta <- read.csv('./data/hyperspectral_imaging/HSI_side_for_biotraits_mod_W9.csv', header = T, sep = ',', dec = '.')
inVar <- "N"
file_path <- "./results/relieF_pca_plsr/Nitrogen/"
```

```{r, echo = F, fig.align='center', fig.width = 4, fig.height = 4}
trait <- phyraw %>%
  select("ID", all_of(inVar), "Treatment") %>%
  na.omit() 
#table(trait$Treatment)
``` 

## Step 1: Feature selection    

```{r, echo=FALSE, fig.align='center', fig.width = 10, fig.height = 5}
ref <- ref_meta[which(ref_meta$ID %in% trait$ID), -c(746:766)] #drop columns with wv > 2500 and meta data 
ref_trait <- merge(ref, subset(trait, select = -Treatment), by = "ID")
ref_trait_2 <- ref_trait[ ,-1]

ref_trait_expRank <- attrEval(N ~., ref_trait_2, estimator = "RReliefFexpRank", ReliefIterations = nrow(trait)*50)
ref_trait_expRank_long <- data.frame(value = unname(ref_trait_expRank), 
                                       wv = as.numeric(gsub("_", ".", sub("...", "",  names(ref_trait_expRank)))))

critical_value <- ref_trait_expRank_long[order(ref_trait_expRank_long$value, decreasing = T)[300], 1]
ref_trait_expRank_long$select <- ifelse(ref_trait_expRank_long$value > critical_value, "Y", "N")

ref_trait_expRank_order <- ref_trait_expRank_long[order(ref_trait_expRank_long$value, decreasing = T)[1:300], ] 
ref_trait_expRank_export <- ref_trait_expRank_long[order(ref_trait_expRank_long$value, decreasing = T), ] 

col_attr <- c("Y"="red", "N"="darkgray")
ref_trait_expRank_plot <- ggplot() +
  geom_line(aes(x = wv, y = value), data = ref_trait_expRank_long, col = "black", size = 0.5) +
  scale_x_continuous(breaks = seq(400, 2500, by = 400), labels = seq(400, 2500, by = 400)) +
  scale_y_continuous(breaks = seq(-0.1, 0.2, by = 0.1), labels = seq(-0.1, 0.2, by = 0.1),
                     limits = c(-0.13, 0.23)) +
  geom_point(aes(x = wv, y = value, col = select), data = ref_trait_expRank_long, size = .7,
             show.legend = F) +
  scale_color_manual(values=col_attr) +
  labs(x= "\n Wavelength (nm)", y = "Attribute values") +                     
  theme_bw() +
  theme(
    axis.title = element_text(size = 25, face = "bold"),
    axis.text = element_text(size = 25, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
        )
ref_trait_expRank_plot
```
Figure 1: Attribute values of HSI data on Week 9 for N (%). Wavelengths with the top 300 attribute values are considered as important (red).  
```{r, echo=FALSE, eval=FALSE}
ggsave(filename = "./results/relieF_pca_plsr/Nitrogen/Nitrogen_expRank.png", plot = ref_trait_expRank_plot, device="png", width = 30, height = 12, units = "cm", dpi = 400)
write.csv(file = paste0(file_path, "reflectanc_N_expRank_export.csv"), x= ref_trait_expRank_export, row.names = F)
```

```{r, echo=FALSE, fig.align='center', fig.width = 10, fig.height = 5}
att_annotate <- ggplot() +
  geom_line(aes(x = wv, y = value), data = ref_trait_expRank_long, col = "black", size = 0.5) +
  geom_point(aes(x = wv, y = value, col = select), data = ref_trait_expRank_long, size = .7,
             show.legend = F) +
  scale_color_manual(values=col_attr) +
  scale_x_continuous(breaks = seq(400, 2500, by = 400), labels = seq(400, 2500, by = 400)) +
  scale_y_continuous(breaks = seq(-0.1, 0.2, by = 0.1), labels = seq(-0.1, 0.2, by = 0.1),
                     limits = c(-0.13, 0.23)) +
  geom_vline(xintercept = c(1500, 1960, 2080, 2115, 2230, 2300), col = "gray") + #N-H stretch
  geom_rect(aes(xmin=480, ymin = 0.15, xmax= 680, ymax = 0.17), #chlorophyll
               col = "#009E73", fill = rgb(0,0,0,0), size = 1.5) +
  annotate("text", x = 1000, y = 0.16, label = "Chlorophyll", col = "#009E73", size = 10) + 
  geom_rect(aes(xmin=680, ymin = 0.18, xmax= 750, ymax = 0.20), #red-edge
               col = "#CC79A7", fill = rgb(0,0,0,0), size = 1.5) +
  annotate("text", x = 1000, y = 0.20, label = "Red-edge", col = "#CC79A7", size = 10) + 
  geom_rect(aes(xmin=1900, ymin = 0.02, xmax= 2100, ymax = 0.11), 
               col = "#0072B2", fill = rgb(0,0,0,0), size = 1.5) +
  annotate("text", x = 1800, y = 0.05, parse = T, label = paste("H[2]*O"), col = "#0072B2", size = 10) + 
  
  labs(x= "\n Wavelength (nm)", y = "Attribute values") +                     
  theme_bw() +
  theme(
    axis.title = element_text(size = 25, face = "bold"),
    axis.text = element_text(size = 25, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
        )
att_annotate
```
Figure 2: Annotations of attribute values of HSI data on Week 9 for N (%). Gray vertical bars are wavelengths that are responsive to N-H stretch. 
```{r, echo=FALSE, eval = F, fig.align='center', fig.width = 10, fig.height = 4}
att_annotate
ggsave(filename = "./results/relieF_pca_plsr/Nitrogen/Nitrogen_att_annotate.png", plot = att_annotate, device="png", width = 30, height = 12, units = "cm", dpi = 400)
```

```{r, echo = F, include=FALSE}
ref_feature <- ref[ , which(as.numeric(gsub("_", ".", sub("...", "",  names(ref)))) %in% ref_trait_expRank_order$wv)]
ref_feature2 <- data.frame(ID = ref$ID, ref_feature)
ref_feature2_long <- melt(ref_feature2, id = "ID")
ref_feature2_long$variable.num <- as.numeric(gsub("_", ".", sub("...", "", ref_feature2_long$variable)))
```
\newpage  

## Step 2: PCA  
The reflectances were not scaled.  

```{r, include=FALSE}
ref_pca <- prcomp(ref_feature)
ref_loadings <- ref_pca$rotation
ref_scores <- ref_pca$x
ref_sd <- ref_pca$sdev
ref_var <- (ref_pca$sdev)^2
ref_var_percent <- ref_var/sum(ref_var)*100
sum(ref_var_percent[1:2])
```

```{r, echo = F, warnings = F, fig.align='center', fig.width = 4, fig.height = 4}
#png(filename = paste0(file_path, "HSI_Nitrogen_var_percent.png"), bg = "white", width = 10, height = 10, units = "in", res = 400)
barplot(ref_var_percent, xlab='PC', ylab='Percent Variance', names.arg=1:length(ref_var_percent), las=1, ylim=c(0, max(ref_var_percent)), col='gray')
#dev.off()
```
Figure 3: PCA scree plot. PCs 1-2 account for 91.8% of total variance. The top ten loadings in PC1 and PC2 are wavelengths from 712 to 724 nm and from 748 to 760 nm, respectively.  
```{r, echo = F, eval = F, include=FALSE}
#record the loadings
ref_PC1_loadings <- ref_loadings[order((ref_loadings[ ,1])^2, decreasing = T), 1]
ref_PC2_loadings <- ref_loadings[order((ref_loadings[ ,2])^2, decreasing = T), 2]

ref_top10_loadings <- data.frame(Trait = "N_percentage", PC = rep(c("PC1", "PC2"), each = 10), 
                    ID = c(as.numeric(gsub("_", ".", sub("...", "", names(ref_PC1_loadings[1:10])))), 
                           as.numeric(gsub("_", ".", sub("...", "", names(ref_PC2_loadings[1:10]))))
                           ),
                    loadings = c(as.numeric(ref_PC1_loadings[1:10]),                   
                                 as.numeric(ref_PC2_loadings[1:10])))
ref_top10_loadings_2 <- ref_top10_loadings[!duplicated(ref_top10_loadings[ ,c('ID')]), ]
write.csv(file = paste0(file_path, "ref_N_percentage_top10_loadings.csv"), x= ref_top10_loadings_2, row.names = F)
```
\newpage  

## Step 3: plsr  

```{r, include=FALSE}
`%notin%` <- Negate(`%in%`)
pls::pls.options(plsralg = "oscorespls")
pls::pls.options("plsralg")
opar <- par(no.readonly = T)
```
1. Prepare calibration and validation datasets  
The data were first grouped by treatments. Then, 80% of the data in each treatment were sampled for the calibration dataset.
```{r, echo = F, fig.align='center', fig.width = 9, fig.height = 5}
ref_I <- as.matrix(ref_scores)
plsr_data <- data.frame(trait, inputs_scores = I(ref_I))
set.seed(999)
prop <- 0.8
group_variables <- c("Treatment")
cal.plsr.data <- plsr_data %>% 
        group_by_at(vars(all_of(group_variables))) %>%
        slice(sample(1:n(), prop*n())) %>% 
        data.frame()
val.plsr.data <- plsr_data[!plsr_data$ID %in% cal.plsr.data$ID,]
print(paste("Cal observations: ",dim(cal.plsr.data)[1],sep="")) #70
print(paste("Val observations: ",dim(val.plsr.data)[1],sep="")) #18

cal_hist_plot <- qplot(cal.plsr.data$N, geom="histogram", binwidth =0.54,
                       main = "Cal. (n = 70)", 
                       xlab = "N (%)", ylab = "Count", 
                       fill=I("#0072B2"), col=I("black")) + 
                       theme_bw(base_size = 18)
 
val_hist_plot <- qplot(val.plsr.data$N, geom="histogram", binwidth =0.54,
                       main = "Val. (n = 18)",
                       xlab = "N (%)", ylab = "Count", 
                       fill=I("#0072B2"), col=I("black")) + 
                       theme_bw(base_size = 18)
cal_val_histograms <- grid.arrange(cal_hist_plot, val_hist_plot, ncol=2)
```
Figure 4: Frequency distributions of N (%) in the calibration (left, *n* = 70) and validation (right, *n* = 18) datasets. 

```{r, echo=FALSE, eval=FALSE, fig.align='center', fig.width = 5, fig.height = 5}
cal_hist_plot
ggsave(filename = paste0(file_path, "cal_nitrogen_histogram.png"), plot = cal_hist_plot, device="png", width = 12, height = 12, units = "cm", dpi = 300)
val_hist_plot
ggsave(filename = paste0(file_path, "val_nitrogen_histogram.png"), plot = val_hist_plot, device="png", width = 12, height = 12, units = "cm", dpi = 300)
write.csv(cal.plsr.data, file= paste0(file_path, "Nitrogen_Cal_PLSR_Dataset.csv"), row.names=FALSE)
write.csv(val.plsr.data, file= paste0(file_path, "Nitrogen_Val_PLSR_Dataset.csv"), row.names=FALSE)
```
\newpage 

2. Fit a plsr model    
Model: N = Scores of PCs. The scores of PCs were not scaled. 
The model was cross-validated with leave-one-out method.  
```{r, echo = F, fig.align='center', fig.width = 7, fig.height = 5}
plsr.ini <- plsr(N~inputs_scores, scale=FALSE,
                 validation="LOO",trace=FALSE,data=cal.plsr.data)
#png(filename = paste0(file_path, "Nitrogen_nComps.png"), bg = "white", width = 5, height = 5, units = "in", res = 400)
par(mar = c(6, 7, 4, 2), mgp = c(5, 2, 0))
nComp <- selectNcomp(plsr.ini, method = "onesigma", plot = T, 
                     cex.axis = 2, cex.lab = 2, xlim = c(0, 15), ylim = c(0.20, 0.55), 
                     font.lab = 2, font = 2)
#dev.off()
```
Figure 5: Root mean square error of prediction (RMSEP) as a function of model component number in PLSR models. The dotted black vertical line is the number of components that has the lowest RMSEP. The vertical long-dashed blue line indicates the suggested number of components. The points and error bars are the means and standard errors of RMSEP in the cross validation, respectively.  
\newpage  

3. Evaluate model performance  
 
```{r, echo = F, fig.align='center', fig.width = 12, fig.height = 6}
plsr.out <- plsr(N~inputs_scores, scale=FALSE, ncomp = nComp,
                 validation="LOO",trace=FALSE,data=cal.plsr.data)
fit <- plsr.out$fitted.values[,,nComp]

cal.plsr.output <- data.frame(measured = cal.plsr.data$N,
                              PLSR_Predicted=fit,
                              PLSR_CV_Predicted=as.vector(plsr.out$validation$pred[,,nComp]))
cal.plsr.output <- cal.plsr.output %>%
  mutate(PLSR_CV_Residuals = PLSR_CV_Predicted-measured)

cal.R2 <- round(pls::R2(plsr.out,intercept=F)[[1]][nComp],5)
cal.RMSEP <- round(sqrt(mean(cal.plsr.output$PLSR_CV_Residuals^2)),5)
cal.NRMSEP <-  round(sqrt(mean(cal.plsr.output$PLSR_CV_Residuals^2))/sd(cal.plsr.output$measured),5)

val.plsr.output <- data.frame(measured = val.plsr.data$N,
                              PLSR_Predicted=as.vector(predict(plsr.out, 
                                                               newdata = val.plsr.data, 
                                                               ncomp=nComp, 
                                                               type="response")[,,]))
val.plsr.output <- val.plsr.output %>%
  mutate(PLSR_Residuals = PLSR_Predicted-measured)

val.R2 <- round(pls::R2(plsr.out,newdata=val.plsr.data,intercept=F)[[1]][nComp],5)
val.RMSEP <- round(sqrt(mean(val.plsr.output$PLSR_Residuals^2)),5)
val.NRMSEP <- round(sqrt(mean(val.plsr.output$PLSR_Residuals^2))/sd(val.plsr.output$measured),5)

rng <- range(cal.plsr.output$measured, cal.plsr.output$PLSR_CV_Predicted,
             val.plsr.output$measured, val.plsr.output$PLSR_Predicted)
text_xloc <- seq(round(rng[1]), round(rng[2]), length = 5)[4]
text_y_R2 <- (round(rng[2])-round(rng[1]))/10*4+round(rng[1])
text_y_RMSEP <- (round(rng[2])-round(rng[1]))/10*3+round(rng[1])
text_y_NRMSEP <- (round(rng[2])-round(rng[1]))/10*2+round(rng[1])

cal_scatter_plot <- ggplot(cal.plsr.output, aes(x=measured, y=PLSR_CV_Predicted)) + 
  theme_bw() + geom_point() + 
  geom_abline(intercept = 0, slope = 1, color="dark grey", linetype="dashed", size=1.5) + 
  geom_smooth(method = "lm", formula = y ~ x, se = F) +
  scale_x_continuous(breaks = seq(round(rng[1]), round(rng[2]), length = 5), 
                     labels = seq(round(rng[1]), round(rng[2]), length = 5), 
                     limits = c(rng[1], rng[2]))+
  scale_y_continuous(breaks = seq(round(rng[1]), round(rng[2]), length = 5), 
                     labels = seq(round(rng[1]), round(rng[2]), length = 5), 
                     limits = c(rng[1], rng[2]))+
  labs(x="\n Measured: Nitrogen (%)",
       y="Predicted: Nitrogen (%)") +
  theme(axis.text=element_text(size=25, face = "bold"), legend.position="none",
        axis.title=element_text(size=25, face = "bold"), 
        axis.text.x = element_text(angle = 0,vjust = 0.5)) + 
 annotate("text", size = 8 , x= text_xloc, y = c(text_y_R2, text_y_RMSEP, text_y_NRMSEP), parse = TRUE, 
           label = c(paste("R^{2} == ", round(cal.R2, 2)), paste("RMSEP ==", round(cal.RMSEP, 2)), paste("NRMSEP ==", round(cal.NRMSEP, 2)))) +
  annotate("text", x = 2, y = 3.75, parse = T, label = "Cal. (italic(n) == 70)", size = 8)
```

```{r, echo=FALSE, eval = F, fig.width=5, fig.height=5, fig.align='center'}  
cal_scatter_plot
ggsave(filename = paste0(file_path, "Nitrogen_Cal_lm_Scatterplot.png"), plot = cal_scatter_plot, device="png", width = 15, height = 15, units = "cm", dpi = 400)
```

```{r, echo=FALSE, fig.width=5, fig.height=5, fig.align='center'}
cal_resid_histogram <- ggplot(cal.plsr.output, aes(x=PLSR_CV_Residuals)) +
  geom_histogram(alpha=.5, position="identity", bins = 8, col="black") + 
  geom_vline(xintercept = 0, color="black", 
             linetype="dashed", size=1) + 
  theme_bw(base_size = 18) + 
  theme(legend.position="none",
        axis.text.x = element_text(angle = 0,vjust = 0.5))+
 labs(title = "Cal. (n = 70)", x = "CV residuals", y = "Count")
```

```{r, echo = F, eval = F, fig.align='center', fig.width = 5, fig.height = 5}
cal_resid_histogram
ggsave(filename = paste0(file_path, "Nitrogen_Cal_Residualsplot.png"), plot = cal_resid_histogram, device="png", width = 12, height = 12, units = "cm", dpi = 300)
```

```{r, echo = F, fig.align='center', fig.width = 5, fig.height = 5}
val_scatter_plot <- ggplot(val.plsr.output, aes(x=measured, y=PLSR_Predicted)) + 
  theme_bw() + geom_point() + 
  geom_abline(intercept = 0, slope = 1, color="dark grey", linetype="dashed", size=1.5) + 
  geom_smooth(method = "lm", formula = y ~ x, se = F) +
  scale_x_continuous(breaks = seq(round(rng[1]), round(rng[2]), length = 5), 
                     labels = seq(round(rng[1]), round(rng[2]), length = 5), 
                     limits = c(rng[1], rng[2]))+
  scale_y_continuous(breaks = seq(round(rng[1]), round(rng[2]), length = 5), 
                     labels = seq(round(rng[1]), round(rng[2]), length = 5), 
                     limits = c(rng[1], rng[2]))+
  labs(x="\n Measured: Nitrogen (%)",
       y="Predicted: Nitrogen (%)") +
  theme(axis.text=element_text(size=25, face = "bold"), legend.position="none",
        axis.title=element_text(size=25, face = "bold"), 
        axis.text.x = element_text(angle = 0,vjust = 0.5))+
   annotate("text", size = 8 , x= text_xloc, y = c(text_y_R2, text_y_RMSEP, text_y_NRMSEP), parse = TRUE, label = c(paste("R^{2} == ", round(val.R2, 2)), paste("RMSEP ==", round(val.RMSEP, 2)), paste("NRMSEP ==", round(val.NRMSEP, 2)))) +
   annotate("text", x = 2, y = 3.75, parse = T, label = "Val. (italic(n) == 18)", size = 8 )
```

```{r, echo=FALSE, eval = F, fig.width=5, fig.height=5, fig.align='center'}  
val_scatter_plot
ggsave(filename = paste0(file_path, "Nitrogen_Val_lm_Scatterplot.png"), plot = val_scatter_plot, device="png", width = 15, height = 15, units = "cm", dpi = 400)
```

```{r, echo = F, fig.align='center', fig.width = 5, fig.height = 5}
val_resid_histogram <- ggplot(val.plsr.output, aes(x=PLSR_Residuals)) +
  geom_histogram(alpha=.5, position="identity", bins = 4, col="black") + 
  geom_vline(xintercept = 0, color="black", 
             linetype="dashed", size=1) + theme_bw(base_size = 18) + 
  theme(legend.position="none",
        axis.text.x = element_text(angle = 0, vjust = 0.5)
        )+
  labs(title = "Val. (n = 18)", x = "Residuals", y = "Count")
```

```{r, echo = F, eval = F, fig.align='center', fig.width = 5, fig.height = 5}
val_resid_histogram
ggsave(filename = paste0(file_path, "Nitrogen_Val_Residualsplot.png"), plot = val_resid_histogram, device="png", width = 12, height = 12, units = "cm", dpi = 300)
```

```{r,  echo = F, fig.align='center', fig.width = 12, fig.height = 6}
scatterplots <- grid.arrange(cal_scatter_plot, val_scatter_plot, ncol = 2)
```
Figure 6: Cross-validated values versus measured values (left, *n* = 70) and external validation values versus measured values (right, *n* = 18) of N (%)
```{r, echo = F, fig.align='center', fig.width = 8, fig.height = 4}
histplots <-  grid.arrange(cal_resid_histogram, val_resid_histogram, ncol=2)
```
Figure 7: Residuals from the cross-validated values versus measured values (left, *n* = 70) and residuals from the external validation values versus measured values (right, *n* = 18) of N (%)  