---
title: "N_reliefF_plsr"
date: "`r Sys.Date ()`"
output:
  pdf_document: default
---

```{r, setup, include=FALSE, cache = F}
knitr::opts_chunk$set(echo = T)
knitr::opts_knit$set(root.dir = '/Users/ting15/Dropbox/Ting_Wang/rice_imaging_ms')
#knitr::opts_knit$set(root.dir = 'C:/Users/rebec/Dropbox/Ting_Wang/rice_imaging_ms')
```

```{r, echo = F, include=FALSE}
rm(list = ls())
## Load package 
list.of.packages <- c("dplyr", "reshape2", "ggplot2", "pls", "scales", "kableExtra", "ggpubr", "gridExtra", "CORElearn", "spectratrait")
invisible(lapply(list.of.packages, library, character.only = TRUE))
```
# Aim  

Quantify leaf-level nitrogen (%, N) at mid-tillering stage with hyperspectral imaging data (HSI data).    

# Methods  

1. Use RReliefFexpRank algorithm to select 300 wavelengths which had the top attributes values for PLSR modeling. These wavelengths were referred to as $W_{PLSR}$.    
2. Build a PLSR model with $W_{PLSR}$ as input and N as output.  

```{r import data, include=FALSE}
phyraw <- read.csv('./data/ground_truth/phyraw_forHSI_2022.csv', header = T, sep = ',', dec = '.')
ref_meta <- read.csv('./data/hyperspectral_imaging/HSI_side_for_biotraits_mod_W9.csv', header = T, sep = ',', dec = '.')
inVar <- "N"
file_path <- "./results/reliefF_plsr/Nitrogen/"
```

```{r select nitrogen as trait, echo = F, fig.align='center', fig.width = 4, fig.height = 4}
trait <- phyraw %>%
  select("ID", all_of(inVar), "Treatment") %>%
  na.omit() 
``` 

## Step 1: Feature selection    

```{r feature selection, echo=FALSE, fig.align='center', fig.width = 10, fig.height = 5}
ref <- ref_meta[which(ref_meta$ID %in% trait$ID), -c(746:766)] #drop columns with wv > 2500 nm and meta data 
ref_trait <- merge(ref, subset(trait, select = -Treatment), by = "ID")
ref_trait_num <- ref_trait[ ,-1] # drop ID

ref_trait_expRank <- attrEval(N ~., ref_trait_num, estimator = "RReliefFexpRank", ReliefIterations = nrow(trait)*50)
ref_trait_expRank_long <- data.frame(value = unname(ref_trait_expRank), 
                                       wv = as.numeric(gsub("_", ".", sub("...", "",  names(ref_trait_expRank)))))

critical_value <- ref_trait_expRank_long[order(ref_trait_expRank_long$value, decreasing = T)[300], 1]
ref_trait_expRank_long$select <- ifelse(ref_trait_expRank_long$value > critical_value, "Y", "N")

ref_trait_expRank_order <- ref_trait_expRank_long[order(ref_trait_expRank_long$value, decreasing = T)[1:300], ] 
ref_trait_expRank_export <- ref_trait_expRank_long[order(ref_trait_expRank_long$value, decreasing = T), ] 

col_attr <- c("Y"="red", "N"="darkgray")
ref_trait_expRank_plot <- ggplot() +
  geom_line(aes(x = wv, y = value), data = ref_trait_expRank_long, col = "black", size = 0.5) +
  scale_x_continuous(breaks = seq(400, 2500, by = 400), labels = seq(400, 2500, by = 400)) +
  scale_y_continuous(breaks = seq(-0.1, 0.2, by = 0.1), labels = seq(-0.1, 0.2, by = 0.1),
                     limits = c(-0.13, 0.23)) +
  geom_point(aes(x = wv, y = value, col = select), data = ref_trait_expRank_long, size = .7,
             show.legend = F) +
  scale_color_manual(values=col_attr) +
  labs(x= "Wavelength (nm)", y = "Attributes values") +                     
  theme_bw() +
  theme(
    axis.title = element_text(size = 25, face = "bold"),
    axis.text = element_text(size = 25, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
        )
ref_trait_expRank_plot
```
Figure 1: Attributes values of HSI data on Week 9 for N (%). Wavelengths with the top 300 attribute values are considered as important (red, $W_{PLSR}$) and are model inputs.  
```{r annotate aattributes values plot, echo=FALSE, fig.align='center', fig.width = 10, fig.height = 5}
att_annotate <- ggplot() +
  geom_line(aes(x = wv, y = value), data = ref_trait_expRank_long, col = "black", size = 0.5) +
  geom_point(aes(x = wv, y = value, col = select), data = ref_trait_expRank_long, size = .7,
             show.legend = F) +
  scale_color_manual(values=col_attr) +
  scale_x_continuous(breaks = seq(400, 2500, by = 400), labels = seq(400, 2500, by = 400)) +
  scale_y_continuous(breaks = seq(-0.1, 0.2, by = 0.1), labels = seq(-0.1, 0.2, by = 0.1),
                     limits = c(-0.13, 0.23)) +
  geom_vline(xintercept = c(1500, 1960, 2080, 2115, 2230, 2300), col = "gray") + #N-H stretch
  geom_rect(aes(xmin=480, ymin = 0.15, xmax= 680, ymax = 0.17), #chlorophyll
               col = "#009E73", fill = rgb(0,0,0,0), size = 1.5) +
  annotate("text", x = 1000, y = 0.16, label = "Chlorophyll", col = "#009E73", size = 10) + 
  geom_rect(aes(xmin=680, ymin = 0.18, xmax= 750, ymax = 0.20), #red-edge
               col = "#CC79A7", fill = rgb(0,0,0,0), size = 1.5) +
  annotate("text", x = 1000, y = 0.20, label = "Red-edge", col = "#CC79A7", size = 10) + 
  geom_rect(aes(xmin=1900, ymin = 0.02, xmax= 2100, ymax = 0.11), 
               col = "#0072B2", fill = rgb(0,0,0,0), size = 1.5) +
  annotate("text", x = 1800, y = 0.05, parse = T, label = paste("H[2]*O"), col = "#0072B2", size = 10) + 
  
  labs(x= "Wavelength (nm)", y = "Attributes values") +                     
  theme_bw() +
  theme(
    axis.title = element_text(size = 25, face = "bold"),
    axis.text = element_text(colour = "black", size = 25, face = "bold"),
    axis.text.y = element_text(angle = 90, hjust = .03),
    panel.border = element_rect(colour = "black", fill=NA, size=2),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
        )
att_annotate
```
```{r att anotate output, echo=FALSE, eval = F, fig.align='center', fig.width = 10, fig.height = 4}
att_annotate
ggsave(filename = paste0(file_path, "Nitrogen_att_annotate.png"), plot = att_annotate, device="png", width = 30, height = 12, units = "cm", dpi = 400)
```
## Step 2: plsr  

Method for pls: Orthogonal scores PLSR  

```{r plsr setting, include=FALSE}
`%notin%` <- Negate(`%in%`)
pls::pls.options(plsralg = "oscorespls")
pls::pls.options("plsralg")
opar <- par(no.readonly = T)
```
1. Prepare calibration and evaluation datasets  
The data were first grouped by treatment. Then, 80% of the data were sampled for the calibration dataset.
```{r, echo = F, fig.align='center', fig.width = 9, fig.height = 5}
ref_select <- as.matrix(ref[ ,names(ref_trait_expRank[ref_trait_expRank >= critical_value])])
plsr_data <- data.frame(trait, Spectra = I(ref_select))
set.seed(999)
prop <- 0.8
group_variables <- c("Treatment")
cal.plsr.data <- plsr_data %>% 
        group_by_at(vars(all_of(group_variables))) %>%
        slice(sample(1:n(), prop*n())) %>% 
        data.frame()
val.plsr.data <- plsr_data[!plsr_data$ID %in% cal.plsr.data$ID,]
print(paste("Cal observations: ",dim(cal.plsr.data)[1],sep="")) #70
print(paste("Val observations: ",dim(val.plsr.data)[1],sep="")) #18

#range(trait$N)
cal_hist_plot <-ggplot(aes(x = N), data = cal.plsr.data) +
  geom_histogram(binwidth =0.3, fill=I("#0072B2"), col=I("black")) +
  coord_cartesian(xlim = c(1,4)) +
  theme_bw(base_size = 18) +
  labs(x = "N (%)", y = "Count") +
  ggtitle("Cal. (n = 70)") 

val_hist_plot <- ggplot(aes(x = N), data = val.plsr.data) +
  geom_histogram(binwidth =0.28, fill=I("#0072B2"), col=I("black")) +
  coord_cartesian(xlim = c(1,4)) +
  theme_bw(base_size = 18) +
  labs(x = "N (%)", y = "Count") +
  ggtitle("Val. (n = 18)")
 
cal_val_histograms <- ggarrange(cal_hist_plot, val_hist_plot, ncol=2)
cal_val_histograms 
```
Figure 2: Frequency distributions of N (%) in the calibration (left, *n* = 70) and validation (right, *n* = 18) datasets. 

2. Fit a plsr model    
 
The model was cross-validated with leave-one-out method.  
```{r determine number of components, echo = F, fig.align='center', fig.width = 7, fig.height = 5}
plsr.ini <- plsr(N~Spectra, scale=FALSE,
                 validation="LOO",trace=FALSE,data=cal.plsr.data)
#png(filename = paste0(file_path, "Nitrogen_nComps.png"), bg = "white", width = 7, height = 5, units = "in", res = 400)
par(mfrow=c(1,1), mar=c(6,7,1,0.5), oma=c(0, 0.1, 0, 0.2), mgp = c(3, 1, 0))
nComps <- selectNcomp(plsr.ini, method = "onesigma", plot = T, lwd=1.6,
                      xlim = c(0, 15), ylim = c(0.20, 0.55), 
                      pch=21, col="black", bg = "grey70",
                      cex.axis = 2, cex.lab = 2, font.lab = 2, font = 2) 
box(lwd=2.2)
#dev.off()
```
Figure 3: Root mean square error of prediction (RMSEP) as a function of model component number in PLSR models. The dotted black vertical line is the number of components that has the lowest RMSEP. The vertical long-dashed blue line indicates the suggested number of components. The points and error bars are the means and standard errors of RMSEP in the cross validation, respectively. 

3. Evaluate model performance  
 
```{r model metrics, echo = F}
plsr.out <- plsr(N~Spectra, scale=FALSE, ncomp = nComps,
                 validation="LOO",trace=FALSE,data=cal.plsr.data)
fit <- plsr.out$fitted.values[,,nComps]
#calibration
cal.plsr.output <- data.frame(N = cal.plsr.data$N,
                              PLSR_Predicted=fit,
                              PLSR_CV_Predicted=as.vector(plsr.out$validation$pred[,,nComps]))
cal.plsr.output <- cal.plsr.output %>%
  mutate(PLSR_CV_Residuals = PLSR_CV_Predicted-N)
#head(cal.plsr.output)
cal.R2 <- round(pls::R2(plsr.out,intercept=F)[[1]][nComps],3) #intercept = F, estimates for a model with zero component does not need to be returned
cal.RMSEP <- round(sqrt(mean(cal.plsr.output$PLSR_CV_Residuals^2)),3)
cal_N_range <- (max(cal.plsr.output$N)-min(cal.plsr.output$N))
cal.NRMSEP <-  round((sqrt(mean(cal.plsr.output$PLSR_CV_Residuals^2))/cal_N_range)*100,3)

val.plsr.output <- data.frame(N = val.plsr.data$N,
                              PLSR_Predicted=as.vector(predict(plsr.out, 
                                                               newdata = val.plsr.data, 
                                                               ncomp=nComps, 
                                                               type="response")[,,]))
val.plsr.output <- val.plsr.output %>%
  mutate(PLSR_Residuals = PLSR_Predicted-N)

val.R2 <- round(pls::R2(plsr.out,newdata=val.plsr.data,intercept=F)[[1]][nComps],3)
val.RMSEP <- round(sqrt(mean(val.plsr.output$PLSR_Residuals^2)),3)
val_N_range <- (max(val.plsr.output$N)-min(val.plsr.output$N))
val.NRMSEP <- round((sqrt(mean(val.plsr.output$PLSR_Residuals^2))/val_N_range)*100,3)
```

```{r cal scatter, echo = F, fig.align='center', fig.width = 9, fig.height = 5}
rng <- range(cal.plsr.output$N, cal.plsr.output$PLSR_CV_Predicted,
             val.plsr.output$N, val.plsr.output$PLSR_Predicted)
text_xloc <- seq(round(rng[1]), round(rng[2]), length = 5)[4]
text_y_R2 <- (round(rng[2])-round(rng[1]))/10*4+round(rng[1])
text_y_RMSEP <- (round(rng[2])-round(rng[1]))/10*3+round(rng[1])
text_y_NRMSEP <- (round(rng[2])-round(rng[1]))/10*2+round(rng[1])

cal_scatter_plot <- ggplot(cal.plsr.output, aes(x=N, y=PLSR_CV_Predicted)) + 
  theme_bw() + geom_point() + 
  geom_abline(intercept = 0, slope = 1, color="dark grey", linetype="dashed", size=1.5) + 
  geom_smooth(method = "lm", formula = y ~ x, se = F) +
  scale_x_continuous(breaks = seq(round(rng[1]), round(rng[2]), length = 5), 
                     labels = seq(round(rng[1]), round(rng[2]), length = 5), 
                     limits = c(rng[1], rng[2]))+
  scale_y_continuous(breaks = seq(round(rng[1]), round(rng[2]), length = 5), 
                     labels = seq(round(rng[1]), round(rng[2]), length = 5), 
                     limits = c(rng[1], rng[2]))+
  labs(x="\n Measured: Nitrogen (%)",
       y="Predicted: Nitrogen (%)") +
  theme(axis.text=element_text(size=25, face = "bold"), legend.position="none",
        axis.title=element_text(size=25, face = "bold"), 
        axis.text.x = element_text(angle = 0,vjust = 0.5)) + 
 annotate("text", size = 8 , x= text_xloc, y = c(text_y_R2, text_y_RMSEP), parse = TRUE, 
           label = c(paste("R^{2} == ", round(cal.R2, 2)), paste("RMSEP ==", round(cal.RMSEP, 2)))) +
  annotate("text", size = 8 , x= text_xloc, y = text_y_NRMSEP, label = "%RMSEP = 13.15") +
  annotate("text", x = 2, y = 3.75, parse = T, label = "Cal. (italic(n) == 70)", size = 8)
```

```{r export cal scatter, echo=FALSE, eval = F, fig.width=5, fig.height=5, fig.align='center'}  
cal_scatter_plot
ggsave(filename = paste0(file_path, "Nitrogen_Cal_lm_Scatterplot.png"), plot = cal_scatter_plot, device="png", width = 15, height = 15, units = "cm", dpi = 400)
```

```{r cal resid, echo=FALSE, fig.width=5, fig.height=5, fig.align='center'}
cal_resid <- ggplot(cal.plsr.output, aes(x = N, y = PLSR_CV_Residuals)) +
  geom_point() + 
  geom_hline(yintercept = 0, color="black", 
             linetype="dashed", size=1) + 
  theme_bw(base_size = 18) +
 labs(title = "Cal. (n = 70)", x = "\n Measured: Nitrogen (%)", y = "CV residuals")
```

```{r export cal resid, echo = F, eval = F, fig.align='center', fig.width = 5, fig.height = 5}
cal_resid
ggsave(filename = paste0(file_path, "Nitrogen_Cal_Residuals.png"), plot = cal_resid, device="png", width = 12, height = 12, units = "cm", dpi = 300)
```

```{r val scatter, echo = F, fig.align='center', fig.width = 5, fig.height = 5}
val_scatter_plot <- ggplot(val.plsr.output, aes(x=N, y=PLSR_Predicted)) + 
  theme_bw() + geom_point() + 
  geom_abline(intercept = 0, slope = 1, color="dark grey", linetype="dashed", size=1.5) + 
  geom_smooth(method = "lm", formula = y ~ x, se = F) +
  scale_x_continuous(breaks = seq(round(rng[1]), round(rng[2]), length = 5), 
                     labels = seq(round(rng[1]), round(rng[2]), length = 5), 
                     limits = c(rng[1], rng[2]))+
  scale_y_continuous(breaks = seq(round(rng[1]), round(rng[2]), length = 5), 
                     labels = seq(round(rng[1]), round(rng[2]), length = 5), 
                     limits = c(rng[1], rng[2]))+
  labs(x="\n Measured: Nitrogen (%)",
       y="Predicted: Nitrogen (%)") +
  theme(axis.text=element_text(size=25, face = "bold"), legend.position="none",
        axis.title=element_text(size=25, face = "bold"), 
        axis.text.x = element_text(angle = 0,vjust = 0.5))+
   annotate("text", size = 8 , x= text_xloc, y = c(text_y_R2, text_y_RMSEP), parse = TRUE, label = c(paste("R^{2} == ", round(val.R2, 2)), paste("RMSEP ==", round(val.RMSEP, 2)))) +
   annotate("text", size = 8 , x= text_xloc, y = text_y_NRMSEP, label = "%RMSEP = 12.63") +
   annotate("text", x = 2, y = 3.75, parse = T, label = "Val. (italic(n) == 18)", size = 8 )
```

```{r export val scatter, echo=FALSE, eval = F, fig.width=5, fig.height=5, fig.align='center'}  
val_scatter_plot
ggsave(filename = paste0(file_path, "Nitrogen_Val_lm_Scatterplot.png"), plot = val_scatter_plot, device="png", width = 15, height = 15, units = "cm", dpi = 400)
```

```{r val resid, echo = F, fig.align='center', fig.width = 5, fig.height = 5}
val_resid <- ggplot(val.plsr.output, aes(x = N, y=PLSR_Residuals)) +
  geom_point() + 
  geom_hline(yintercept = 0, color="black", 
             linetype="dashed", size=1) + theme_bw(base_size = 18) + 
  labs(title = "Val. (n = 18)", x = "\n Measured: Nitrogen (%)", y = "Residuals")
```

```{r export val resid, echo = F, eval = F, fig.align='center', fig.width = 5, fig.height = 5}
val_resid
ggsave(filename = paste0(file_path, "Nitrogen_Val_Residuals.png"), plot = val_resid, device="png", width = 12, height = 12, units = "cm", dpi = 300)
```

```{r scatterplots,  echo = F, fig.align='center', fig.width = 12, fig.height = 6}
scatterplots <- grid.arrange(cal_scatter_plot, val_scatter_plot, ncol = 2)
```
Figure 4: Cross-validated values versus measured values (left, *n* = 70) and external validation values versus measured values (right, *n* = 18) of N (%)
```{r resid plots, echo = F, fig.align='center', fig.width = 8, fig.height = 4}
residsplots <-  grid.arrange(cal_resid, val_resid, ncol=2)
```
Figure 5: Residuals from the cross-validated values versus measured values (left, *n* = 70) and residuals from the external validation values versus measured values (right, *n* = 18) of N (%)  
```{r Jackknife validation, echo=FALSE}
if(grepl("Windows", sessionInfo()$running)){
  pls.options(parallel =NULL)
} else {
  pls.options(parallel = parallel::detectCores()-1)
}

seg <- 7
jk.plsr.out <- pls::plsr(as.formula(paste(inVar,"~","Spectra")), scale=FALSE, 
                         center=TRUE, ncomp=nComps, 
                         validation="CV", segments = seg, 
                         segment.type="interleaved", trace=FALSE, 
                         jackknife=TRUE, data=cal.plsr.data)
pls.options(parallel = NULL)

Jackknife_coef <- spectratrait::f.coef.valid(plsr.out = jk.plsr.out, data_plsr = cal.plsr.data, 
                               ncomp = nComps, inVar=inVar)
Jackknife_intercept <- Jackknife_coef[1,,,]
Jackknife_coef <- Jackknife_coef[2:dim(Jackknife_coef)[1],,,]

interval <- c(0.025,0.975)
Jackknife_Pred <- val.plsr.data$Spectra %*% Jackknife_coef + 
  matrix(rep(Jackknife_intercept, length(val.plsr.data[,inVar])), byrow=TRUE, 
         ncol=length(Jackknife_intercept))
Interval_Conf <- apply(X = Jackknife_Pred,MARGIN = 1,
                       FUN = quantile,probs=c(interval[1],interval[2]))
sd_mean <- apply(X = Jackknife_Pred,MARGIN = 1,FUN =sd)
sd_res <- sd(val.plsr.output$PLSR_Residuals)
sd_tot <- sqrt(sd_mean^2+sd_res^2)
val.plsr.output$LCI <- Interval_Conf[1,]
val.plsr.output$UCI <- Interval_Conf[2,]
val.plsr.output$LPI <- val.plsr.output$PLSR_Predicted-1.96*sd_tot
val.plsr.output$UPI <- val.plsr.output$PLSR_Predicted+1.96*sd_tot
#head(val.plsr.output)
```

```{r valid pred intervals, Jackknife validation plot, echo=FALSE, fig.align='center', fig.height=12, fig.width=12}
val.plsr.output$Treatment <- val.plsr.data$Treatment
val.plsr.output.N1 <- val.plsr.output %>%
  filter(val.plsr.output$Treatment == "N1")
val.plsr.output.N2 <- val.plsr.output %>%
  filter(val.plsr.output$Treatment == "N2")
expr <- vector("expression", 3)
expr[[1]] <- bquote(R^2==.(val.R2))
expr[[2]] <- bquote(RMSEP==.(val.RMSEP))
expr[[3]] <- bquote("%RMSEP"==.(val.NRMSEP))
rng_vals <- c(min(val.plsr.output$LPI), max(val.plsr.output$UPI))
par(mfrow=c(1,1), mar=c(6, 7, 1, 0.4), oma=c(0, 0.1, 0, 0.2), mgp = c(5, 2, 0)) 
plotrix::plotCI(val.plsr.output.N1[,"N"], val.plsr.output.N1$PLSR_Predicted, 
       li=val.plsr.output.N1$LPI, ui=val.plsr.output.N1$UPI, gap=0.009,sfrac=0.004, 
       lwd=1.6, xlim=c(rng_vals[1], rng_vals[2]), ylim=c(rng_vals[1], rng_vals[2]), 
       err="y", pch=21, col="black", pt.bg="olivedrab4", scol="olivedrab3",
       cex=2,ylab=paste0("Predicted ", paste(inVar), " (%)"),
       xlab=paste0("Measured ", paste(inVar), " (%)"),
       cex.axis=2, cex.lab = 3, font.lab = 2, font = 2)
plotrix::plotCI(val.plsr.output.N2[,"N"], val.plsr.output.N2$PLSR_Predicted, add = T,
       li=val.plsr.output.N2$LPI, ui=val.plsr.output.N2$UPI, gap=0.009,sfrac=0.004, 
       lwd=1.6, xlim=c(rng_vals[1], rng_vals[2]), ylim=c(rng_vals[1], rng_vals[2]), 
       err="y", pch=21, col="black", pt.bg="orange3", scol="orange2", cex=2)
abline(0,1,lty=2,lw=2)
legend(rng_vals[1], rng_vals[2], legend=expr, bty="n", cex=2, y.intersp = 1.5)
legend("bottomright", inset = c(0.1,0.1), legend = c("N1", "N2"), pch=21 , pt.bg = c("olivedrab4", "orange3"),
       col = "black", cex=2, pt.cex = 2, text.col = "black", y.intersp = 1.5, bty = "n")
box(lwd=2.2)
dev.copy(png, file.path(paste0(file_path, "N_model_performance.png")), height=2800, width=3200, res=340)
dev.off()

```

```{r cal plot with aesthetic from model performance plot, echo=FALSE, fig.align='center', fig.height=12, fig.width=12}
expr.cal <- vector("expression", 3)
expr.cal[[1]] <- bquote(R^2==.(cal.R2))
expr.cal[[2]] <- bquote(RMSEP==.(cal.RMSEP))
expr.cal[[3]] <- bquote("%RMSEP"==.(cal.NRMSEP))
cal.plsr.output$Treatment <- cal.plsr.data$Treatment

cal.plsr.output.N1 <- cal.plsr.output %>%
  filter(cal.plsr.output$Treatment == "N1")
cal.plsr.output.N2 <- cal.plsr.output %>%
  filter(cal.plsr.output$Treatment == "N2")

par(mfrow=c(1,1), mar=c(6, 7, 1, 0.4), oma=c(0, 0.1, 0, 0.2), mgp = c(5, 2, 0))
plot(cal.plsr.output.N1[,"N"], cal.plsr.output.N1$PLSR_CV_Predicted, 
       lwd=1.6, xlim=c(rng_vals[1], rng_vals[2]), ylim=c(rng_vals[1], rng_vals[2]), 
       pch=21, col="black", bg = "olivedrab4",
       cex=2, ylab=paste0("Predicted ", paste(inVar), " (%)"),
       xlab=paste0("Measured ", paste(inVar), " (%)"),
       cex.axis=2, cex.lab = 3, font.lab = 2, font = 2)
points(cal.plsr.output.N2[,"N"], cal.plsr.output.N2$PLSR_CV_Predicted,
       pch=21, col="black", bg = "orange3", cex=2)
abline(0,1,lty=2,lw=2)
legend(rng_vals[1], rng_vals[2], legend=expr.cal, bty="n", cex=2, y.intersp = 1.5)
box(lwd=2.2)
dev.copy(png, file.path(paste0(file_path, "N_Cal_aes_w_model_performance.png")), height=2800, width=3200, res=340)
dev.off()

```